---
title: "Exploratory Analysis"
author: "Edward Bosko"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=F, message=F}
library(tidyverse)
library(readxl)
library(MASS)
library(ggplot2)
library(brant)
library(emmeans)
```

## Purpose

This file will be used for exploratory analysis: evaluating the spread and distribution of the data, checking proportional odds assumption of simple models without random effects, generating summary statistics of demographic data

## Reading in Data

```{r, warning=F, message=F}
#read in data
tib <- read_excel('C:/Users/Eddie/OneDrive/Desktop/BIOSTAT/Thesis/Thesis/DataProcessed/2025.4.2.xlsx', sheet = "tib.long")

ggo <- read_excel('C:/Users/Eddie/OneDrive/Desktop/BIOSTAT/Thesis/Thesis/DataProcessed/2025.4.2.xlsx', sheet = "ggo.long")

cons <- read_excel('C:/Users/Eddie/OneDrive/Desktop/BIOSTAT/Thesis/Thesis/DataProcessed/2025.4.2.xlsx', sheet = "cons.long")

bronch <- read_excel('C:/Users/Eddie/OneDrive/Desktop/BIOSTAT/Thesis/Thesis/DataProcessed/2025.4.2.xlsx', sheet = "bronch.long")

atel <- read_excel('C:/Users/Eddie/OneDrive/Desktop/BIOSTAT/Thesis/Thesis/DataProcessed/2025.4.2.xlsx', sheet = "atel.long")

ln <- read_excel('C:/Users/Eddie/OneDrive/Desktop/BIOSTAT/Thesis/Thesis/DataProcessed/2025.4.2.xlsx', sheet = "ln.long")

thin <- read_excel('C:/Users/Eddie/OneDrive/Desktop/BIOSTAT/Thesis/Thesis/DataProcessed/2025.4.2.xlsx', sheet = "thin.long")

thick <- read_excel('C:/Users/Eddie/OneDrive/Desktop/BIOSTAT/Thesis/Thesis/DataProcessed/2025.4.2.xlsx', sheet = "thick.long")

demo <- read.csv('C:/Users/Eddie/OneDrive/Desktop/BIOSTAT/Thesis/Thesis/DataRaw/demographic_final.csv')

flat <- read.csv('C:/Users/Eddie/OneDrive/Desktop/BIOSTAT/Thesis/Thesis/DataRaw/demographic_final.csv')
```

## Visualize the Distribution of Scores

### Tree-in-bud

```{r}
score_colors_new <- c(
  "3" = "#E34A33",  # Orange-Red for Score 3
  "2" = "#3B9142",  # Green for Score 2
  "1" = "#5EC9C2",  # Light Blue for Score 1
  "0" = "white"   # white for Score 0
)

#by rating 
tib %>%
  group_by(rater, score) %>%
  summarize(n=n()) %>%
  mutate(prop = n / sum(n))

ggplot(tib, aes(x=factor(score), fill=rater)) +
  geom_bar(position = 'dodge') +
  labs(title = 'Distribution of Tree-in-bud Ratings',
       x='Rating',
       y ='Count')

#by lobe
tib %>%
  group_by(lobe, score) %>%
  summarize(n=n()) %>%
  mutate(prop = n / sum(n))

tib %>%
  mutate(score = factor(score, levels = c(3,2,1,0)),
         lobe = factor(lobe, levels = c('RUL','RML','RLL','LUS','LLS','LLL'))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack', color = 'black') + 
  scale_fill_manual(values = score_colors_new) +
  labs(
    title = 'Distribution of TIB Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()

tib %>%
  filter(rater == 'JW') %>%
  mutate(score = factor(score, levels = c(3,2,1,0))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack') + 
  labs(
    title = 'Distribution of JW\'s TIB Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()

tib %>%
  filter(rater == 'VH') %>%
  mutate(score = factor(score, levels = c(3,2,1,0))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack') + 
  labs(
    title = 'Distribution of VH\'s TIB Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()
```

### Large Nodules

```{r}
#by rating 
ln %>%
  group_by(rater, score) %>%
  summarize(n=n()) %>%
  mutate(prop = n / sum(n))

ggplot(ln, aes(x=factor(score), fill=rater)) +
  geom_bar(position = 'dodge') +
  labs(title = 'Distribution of Large Nodule Ratings',
       x='Rating',
       y ='Count')

#by lobe
ln %>%
  group_by(lobe, score) %>%
  summarize(n=n()) %>%
  mutate(prop = n / sum(n))

ln %>%
  mutate(score = factor(score, levels = c(3,2,1,0)),
         lobe = factor(lobe, levels = c('RUL','RML','RLL','LUS','LLS','LLL'))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack') + 
  labs(
    title = 'Distribution of LN Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()

ln %>%
  filter(rater == 'JW') %>%
  mutate(score = factor(score, levels = c(3,2,1,0))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack') + 
  labs(
    title = 'Distribution of JW\'s LN Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()

ln %>%
  filter(rater == 'VH') %>%
  mutate(score = factor(score, levels = c(3,2,1,0))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack') + 
  labs(
    title = 'Distribution of VH\'s LN Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()
```

### Ground-glass Opacities

```{r}
#by rating 
ggo %>%
  group_by(rater, score) %>%
  summarize(n=n()) %>%
  mutate(prop = n / sum(n))

ggplot(ggo, aes(x=factor(score), fill=rater)) +
  geom_bar(position = 'dodge') +
  labs(title = 'Distribution of Ground Glass Opacity Ratings',
       x='Rating',
       y ='Count')

#by lobe
ggo %>%
  group_by(lobe, score) %>%
  summarize(n=n()) %>%
  mutate(prop = n / sum(n))

ggo %>%
  mutate(score = factor(score, levels = c(3,2,1,0)),
         lobe = factor(lobe, levels = c('RUL','RML','RLL','LUS','LLS','LLL'))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack', color = 'black') + 
  scale_fill_manual(values = score_colors_new) +
  labs(
    title = 'Distribution of GGO Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()

ggo %>%
  filter(rater == 'JW') %>%
  mutate(score = factor(score, levels = c(3,2,1,0))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack') + 
  labs(
    title = 'Distribution of JW\'s GGO Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()

ggo %>%
  filter(rater == 'VH') %>%
  mutate(score = factor(score, levels = c(3,2,1,0))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack') + 
  labs(
    title = 'Distribution of VH\'s GGO Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()
```

### Consolidations

```{r}
#by rating 
cons %>%
  group_by(rater, score) %>%
  summarize(n=n()) %>%
  mutate(prop = n / sum(n))

ggplot(cons, aes(x=factor(score), fill=rater)) +
  geom_bar(position = 'dodge') +
  labs(title = 'Distribution of Consolidation Ratings',
       x='Rating',
       y ='Count')

#by lobe
cons %>%
  group_by(lobe, score) %>%
  summarize(n=n()) %>%
  mutate(prop = n / sum(n))

cons %>%
  mutate(score = factor(score, levels = c(3,2,1,0)),
         lobe = factor(lobe, levels = c('RUL','RML','RLL','LUS','LLS','LLL'))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack', color = 'black') + 
  scale_fill_manual(values = score_colors_new) +
  labs(
    title = 'Distribution of Consolidation Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()

cons %>%
  filter(rater == 'JW') %>%
  mutate(score = factor(score, levels = c(3,2,1,0))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack') + 
  labs(
    title = 'Distribution of JW\'s Consolidation Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()

cons %>%
  filter(rater == 'VH') %>%
  mutate(score = factor(score, levels = c(3,2,1,0))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack') + 
  labs(
    title = 'Distribution of VH\'s Consolidation Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()
```

### Bronchiectasis

```{r}
#by rating 
bronch %>%
  group_by(rater, score) %>%
  summarize(n=n()) %>%
  mutate(prop = n / sum(n))

ggplot(bronch, aes(x=factor(score), fill=rater)) +
  geom_bar(position = 'dodge') +
  labs(title = 'Distribution of Bronchiectasis Ratings',
       x='Rating',
       y ='Count')

#by lobe
bronch %>%
  group_by(lobe, score) %>%
  summarize(n=n()) %>%
  mutate(prop = n / sum(n))

bronch %>%
  mutate(score = factor(score, levels = c(3,2,1,0)),
         lobe = factor(lobe, levels = c('RUL','RML','RLL','LUS','LLS','LLL'))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack', color = 'black') + 
  scale_fill_manual(values = score_colors_new) +
  labs(
    title = 'Distribution of Bronchiectasis Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()

bronch %>%
  filter(rater == 'JW') %>%
  mutate(score = factor(score, levels = c(3,2,1,0))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack') + 
  labs(
    title = 'Distribution of JW\'s Bronchiectasis Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()

bronch %>%
  filter(rater == 'VH') %>%
  mutate(score = factor(score, levels = c(3,2,1,0))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack') + 
  labs(
    title = 'Distribution of VH\'s Bronchiectasis Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()
```

### Atelectasis

```{r}
#by rating 
atel %>%
  group_by(rater, score) %>%
  summarize(n=n()) %>%
  mutate(prop = n / sum(n))

ggplot(atel, aes(x=factor(score), fill=rater)) +
  geom_bar(position = 'dodge') +
  labs(title = 'Distribution of Atelectasis Ratings',
       x='Rating',
       y ='Count')

#by lobe
atel %>%
  group_by(lobe, score) %>%
  summarize(n=n()) %>%
  mutate(prop = n / sum(n))

atel %>%
  mutate(score = factor(score, levels = c(3,2,1,0)),
         lobe = factor(lobe, levels = c('RUL','RML','RLL','LUS','LLS','LLL'))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack', color = 'black') + 
  scale_fill_manual(values = score_colors_new) +
  labs(
    title = 'Distribution of Atelectasis Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()

atel %>%
  filter(rater == 'JW') %>%
  mutate(score = factor(score, levels = c(3,2,1,0))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack') + 
  labs(
    title = 'Distribution of JW\'s Atelectasis Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()

atel %>%
  filter(rater == 'VH') %>%
  mutate(score = factor(score, levels = c(3,2,1,0))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack') + 
  labs(
    title = 'Distribution of VH\'s Atelectasis Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()
```

### Thin Wall Cavity

```{r}
#by rating 
thin %>%
  group_by(rater, score) %>%
  summarize(n=n()) %>%
  mutate(prop = n / sum(n))

ggplot(thin, aes(x=factor(score), fill=rater)) +
  geom_bar(position = 'dodge') +
  labs(title = 'Distribution of Thin Wall Cavity Ratings',
       x='Rating',
       y ='Count')

#by lobe
thin %>%
  group_by(lobe, score) %>%
  summarize(n=n()) %>%
  mutate(prop = n / sum(n))

thin %>%
  mutate(score = factor(score, levels = c(3,2,1,0)),
         lobe = factor(lobe, levels = c('RUL','RML','RLL','LUS','LLS','LLL'))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack') + 
  labs(
    title = 'Distribution of Thin Wall Cavity Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()

thin %>%
  filter(rater == 'JW') %>%
  mutate(score = factor(score, levels = c(3,2,1,0))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack') + 
  labs(
    title = 'Distribution of JW\'s Thin Wall Cavity Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()

thin %>%
  filter(rater == 'VH') %>%
  mutate(score = factor(score, levels = c(3,2,1,0))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack') + 
  labs(
    title = 'Distribution of VH\'s Thin Wall Cavity Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()
```

### Thick Wall Cavity

```{r}
#by rating 
thick %>%
  group_by(rater, score) %>%
  summarize(n=n()) %>%
  mutate(prop = n / sum(n))

ggplot(thick, aes(x=factor(score), fill=rater)) +
  geom_bar(position = 'dodge') +
  labs(title = 'Distribution of Thick Wall Cavity Ratings',
       x='Rating',
       y ='Count')

#by lobe
thick %>%
  group_by(lobe, score) %>%
  summarize(n=n()) %>%
  mutate(prop = n / sum(n))

thick %>%
  mutate(score = factor(score, levels = c(3,2,1,0)),
         lobe = factor(lobe, levels = c('RUL','RML','RLL','LUS','LLS','LLL'))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack') + 
  labs(
    title = 'Distribution of Thick Wall Cavity Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()

thick %>%
  filter(rater == 'JW') %>%
  mutate(score = factor(score, levels = c(3,2,1,0))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack') + 
  labs(
    title = 'Distribution of JW\'s Thick Wall Cavity Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()

thick %>%
  filter(rater == 'VH') %>%
  mutate(score = factor(score, levels = c(3,2,1,0))) %>%
  ggplot(aes (x=lobe, fill = factor(score))) +
  geom_bar(position = 'stack') + 
  labs(
    title = 'Distribution of VH\'s Thick Wall Cavity Ratings by Lobe',
    x='Lobe',
    y='Count',
    fill='Rating'
  ) +
  theme_minimal()
```

## Heatmap

### Ordinal Outcomes

```{r}
# Add feature labels
tib$feature <- "TIB"
ggo$feature <- "GGO"
bronch$feature <- "Bronchiectasis"
cons$feature <- "Consolidation"
atel$feature <- "Atelectasis"

# Bind together
ordinal_data <- bind_rows(tib, ggo, bronch, cons, atel)

# Create heatmap data: average score per lobe and feature
heatmap_data_ordinal <- ordinal_data %>%
  group_by(lobe, feature) %>%
  summarize(mean_score = mean(score, na.rm = TRUE)) %>%
  ungroup()

ggplot(heatmap_data_ordinal, aes(x = feature, y = lobe, fill = mean_score)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "lightblue", high = "darkblue", name = "Avg Score") +
  labs(
    title = "Average CT Score by Feature and Lobe (Ordinal Features)",
    x = "CT Feature",
    y = "Lung Lobe"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Binary Outcomes

```{r}
# Add feature labels
ln$feature <- "Large Nodules"
thin$feature <- "Thin-wall Cavity"
thick$feature <- "Thick-wall Cavity"

# Bind
binary_data <- bind_rows(ln, thin, thick)

# Create heatmap data
heatmap_data_binary <- binary_data %>%
  group_by(lobe, feature) %>%
  summarize(mean_score = mean(as.numeric(score), na.rm = TRUE)) %>%
  ungroup()

# Plot heatmap
ggplot(heatmap_data_binary, aes(x = feature, y = lobe, fill = mean_score)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "lightyellow", high = "red", name = "Proportion Present") +
  labs(
    title = "Proportion of Binary CT Feature Present by Lobe",
    x = "CT Feature",
    y = "Lung Lobe"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

## Small Multiples

```{r}
# Add feature label for each ordinal dataset
tib$feature <- "TIB"
ggo$feature <- "GGO"
bronch$feature <- "BRONCH"
cons$feature <- "CONS"
atel$feature <- "ATEL"

# Combine all ordinal features into one
ordinal_all <- bind_rows(tib, ggo, bronch, cons, atel)

# Clean and reformat
ordinal_all <- ordinal_all %>%
  mutate(
    score = as.numeric(as.character(score)),  # ensure numeric
    lobe = factor(lobe, levels = c("RUL", "RML", "RLL", "LUS", "LLS", "LLL")),
    feature = factor(feature, levels = c("TIB", "GGO", "CONS", "ATEL", "BRONCH"))
  )

# Create small multiples matrix of score distributions
ggplot(ordinal_all, aes(x = rater, y = score, fill = rater)) +
  geom_violin(scale = "width", adjust = 1, trim = FALSE, alpha = 0.6, color = NA) +
  geom_boxplot(width = 0.1, outlier.shape = NA, alpha = 0.9) +
  facet_grid(feature ~ lobe) +
  scale_y_continuous(breaks = c(0, 3)) +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = "Score Distributions by Lobe and CT Feature",
    x = "Rater",
    y = "Score"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

```

## Simple Models without Random Effects; Lobe and Rater as Fixed Effects

### Tree-in-bud

```{r}
#convert score to an ordered factor (0 < 1 < 2 < 3)
tib$score <- ordered(tib$score, levels = c(0,1,2,3))

tib$lobe <- factor(tib$lobe, 
                           levels = c("RUL", "RML", "RLL", "LUS", "LLS", "LLL"))

#fit simple PO model, ignoring REs
tib_polr <- polr(
  score ~ lobe + rater, 
  data = tib, 
  method = "logistic", 
  Hess = TRUE
)

#look at summary of results
summary(tib_polr)

#check the proportional odds assumption with the brant test
brant(tib_polr)

print("AIC:")
AIC(tib_polr)

#set up contrasts to compare lobes
tib_emm <- emmeans(tib_polr, ~ lobe)
tib_emm
pairs(tib_emm)
```

Proportional odds violated

Significant differences:
LUS vs all other lobes
*RUL - LUS: OR = 2.19 => RUL's odds of disease is 2.19 times that of LUS
*RML - LUS: OR = 2.45 => RML's odds of disease is 2.45 times that of LUS
*RLL - LUS: OR = 3.14 => RLL's odds of disease is 3.14 times that of LUS
*RLL - LLS: OR = 1.68 => RLL's odds of disease is 1.68 times that of LUS
*LUS - LLS: OR = 0.54 => LUS's odds of disease is 0.54 times that of LLS
*LUS - LLL: OR = 0.37 => LLL's odds of disease is 0.37 times that of LUS

### Large Nodules

```{r}
#ensure score is a factor with levels 0 and 1
ln$score <- factor(ln$score, levels = c(0,1))

#make lobe a factor in the desired order
ln$lobe <- factor(ln$lobe, 
                             levels = c("RUL", "RML", "RLL", "LUS", "LLS", "LLL"))

#fit a standard logistic regression model (binary) 
ln_glm <- glm(score ~ lobe + rater, 
                 data = ln, 
                 family = binomial(link = "logit"))

#look at model summary 
summary(ln_glm)

#obtain estimated means (probabilities of score=1 by lobe)
ln_emm <- emmeans(ln_glm, ~ lobe, type = "response")

#see the pairwise comparisons
pairs(ln_emm)
```

Significant findings:
*RUL - RML: OR = 2.73 => RUL's odds of disease is 2.73 times that of RML
*RUL - LUS: OR = 2.94 => RUL's odds of disease is 2.94 times that of LUS
*RUL - LLS: OR = 2.73 => RUL's odds of disease is 2.73 times that of LLS
*RML - RLL: OR = 0.35 => RML's odds of disease is 0.35 times that of RLL
*RML - LLL: OR = 0.32 => RML's odds of disease is 0.32 times that of LLL
*RLL - LUS: OR = 3.06 => RLL's odds of disease is 3.06 times that of LUS
*RLL - LLS: OR = 2.84 => RLL's odds of disease is 2.84 times that of LLS
*LUS - LLL: OR = 0.30 => LUS's odds of disease is 0.30 times that of LLL
*LLS - LLL: OR = 0.32 => LLS's odds of disease is 0.32 times that of LLL

### Ground-glass Opacities

```{r}
#convert score to an ordered factor (0 < 1 < 2 < 3)
ggo$score <- ordered(ggo$score, levels = c(0,1,2,3))

ggo$lobe <- factor(ggo$lobe, 
                           levels = c("RUL", "RML", "RLL", "LUS", "LLS", "LLL"))

#fit simple PO model, ignoring REs
ggo_polr <- polr(
  score ~ lobe + rater, 
  data = ggo, 
  method = "logistic", 
  Hess = TRUE
)

#look at summary of results
summary(ggo_polr)

#check the proportional odds assumption with the brant test
brant(ggo_polr)

#set up contrasts to compare lobes
ggo_emm <- emmeans(ggo_polr, ~ lobe)
ggo_emm
pairs(ggo_emm)
```

Proportional odds not violated

No significant differences between lobes

### Consolidations

```{r}
#convert score to an ordered factor (0 < 1 < 2 < 3)
cons$score <- ordered(cons$score, levels = c(0,1,2,3))

cons$lobe <- factor(cons$lobe, 
                           levels = c("RUL", "RML", "RLL", "LUS", "LLS", "LLL"))

#fit simple PO model, ignoring REs
cons_polr <- polr(
  score ~ lobe + rater, 
  data = cons, 
  method = "logistic", 
  Hess = TRUE
)

#look at summary of results
summary(cons_polr)

#check the proportional odds assumption with the brant test
brant(cons_polr)

#set up contrasts to compare lobes
cons_emm <- emmeans(cons_polr, ~ lobe)
cons_emm
pairs(cons_emm)
```

Proportional odds assumption not violated

Significant findings:
*RUL - LUS: OR = 3.03 => RUL's odds of disease is 3.03 times that of LUS
*RUL - LLL: OR = 1.74 => RUL's odds of disease is 1.74 times that of LLL
*RML - LUS: OR = 3.38 => RML's odds of disease is 3.38 times that of LUS
*RML - LLL: OR = 1.93 => RML's odds of disease is 1.93 times that of LLL
*RLL - LUS: OR = 2.11 => RLL's odds of disease is 2.11 times that of LUS
*LUS - LLS: OR = 0.35 => LUS's odds of disease is 0.35 times that of LLS

### Bronchiectasis

```{r}
#convert score to an ordered factor (0 < 1 < 2 < 3)
bronch$score <- ordered(bronch$score, levels = c(0,1,2,3))

bronch$lobe <- factor(bronch$lobe, 
                           levels = c("RUL", "RML", "RLL", "LUS", "LLS", "LLL"))

#fit simple PO model, ignoring REs
bronch_polr <- polr(
  score ~ lobe + rater, 
  data = bronch, 
  method = "logistic", 
  Hess = TRUE
)

#look at summary of results
summary(bronch_polr)

#check the proportional odds assumption with the brant test
brant(bronch_polr)

#set up contrasts to compare lobes
bronch_emm <- emmeans(bronch_polr, ~ lobe)
bronch_emm
pairs(bronch_emm)
```

Proportional odds assumption violated

Significant findings:
*RUL - RML: OR = 0.42 => RUL's odds of disease is 0.42 times that of RML
*RUL - LUS: OR = 3.03 => RUL's odds of disease is 3.03 times that of LUS
*RUL - LLL: OR = 1.70 => RUL's odds of disease is 1.70 times that of LLL
*RML - RLL: OR = 2.80 => RML's odds of disease is 2.80 times that of RLL
*RML - LUS: OR = 7.21 => RML's odds of disease is 7.21 times that of LUS
*RML - LLS: OR = 1.68 => RML's odds of disease is 1.68 times that of LLS
*RML - LLL: OR = 4.05 => RML's odds of disease is 4.05 times that of LLL
*RLL - LUS: OR = 2.58 => RLL's odds of disease is 2.58 times that of LUS
*RLL - LLS: OR = 0.60 => RLL's odds of disease is 0.60 times that of LLS
*LUS - LLS: OR = 0.23 => LUS's odds of disease is 0.23 times that of LLS
*LUS - LLL: OR = 0.56 => LUS's odds of disease is 0.56 times that of LLL
*LLS - LLL: OR = 2.41 => LLS's odds of disease is 2.41 times that of LLL

### Atelectasis

```{r}
#convert score to an ordered factor (0 < 1 < 2 < 3)
atel$score <- ordered(atel$score, levels = c(0,1,2,3))

atel$lobe <- factor(atel$lobe, 
                           levels = c("RUL", "RML", "RLL", "LUS", "LLS", "LLL"))

#fit simple PO model, ignoring REs
atel_polr <- polr(
  score ~ lobe + rater, 
  data = atel, 
  method = "logistic", 
  Hess = TRUE
)

#look at summary of results
summary(atel_polr)

#check the proportional odds assumption with the brant test
brant(atel_polr)

#set up contrasts to compare lobes
atel_emm <- emmeans(atel_polr, ~ lobe)
atel_emm
pairs(atel_emm)
```

Proportional odds assumption holds

Significant findings:
*RUL - RML: OR = 0.18 => RUL's odds of disease is 0.22 times that of RML
*RUL - RLL: OR = 2.47 => RUL's odds of disease is 2.47 times that of RLL
*RUL - LUS: OR = 2.43 => RUL's odds of disease is 2.43 times that of LUS
*RUL - LLS: OR = 0.25 => RUL's odds of disease is 0.25 times that of LLS
*RML - RLL: OR = 13.4 => RML's odds of disease is 13.4 times that of RLL
*RML - LUS: OR = 13.2 => RML's odds of disease is 13.2 times that of LUS
*RML - LLL: OR = 10.8 => RML's odds of disease is 10.8 times that of LLL
*RLL - LLS: OR = 0.10 => RLL's odds of disease is 0.10 times that of LLS
*LUS - LLS: OR = 0.10 => LUS's odds of disease is 0.10 times that of LLS
*LLS - LLL: OR = 7.96 => LLS's odds of disease is 7.96 times that of LLL

### Thin Wall Cavity

```{r}
#ensure score is a factor with levels 0 and 1
thin$score <- factor(thin$score, levels = c(0,1))

#make lobe a factor in the desired order
thin$lobe <- factor(thin$lobe, 
                             levels = c("RUL", "RML", "RLL", "LUS", "LLS", "LLL"))

#fit a standard logistic regression model (binary) 
thin_glm <- glm(score ~ lobe + rater, 
                 data = thin, 
                 family = binomial(link = "logit"))

#look at model summary 
summary(thin_glm)

#obtain estimated means (probabilities of score=1 by lobe)
thin_emm <- emmeans(thin_glm, ~ lobe, type = "response")

#see the pairwise comparisons
pairs(thin_emm)
```

Lobes do not significantly differ


### Thick Wall Cavity

```{r}
#ensure score is a factor with levels 0 and 1
thick$score <- factor(thick$score, levels = c(0,1))

#make lobe a factor in the desired order
thick$lobe <- factor(thick$lobe, 
                             levels = c("RUL", "RML", "RLL", "LUS", "LLS", "LLL"))

#fit a standard logistic regression model (binary) 
thick_glm <- glm(score ~ lobe + rater, 
                 data = thick, 
                 family = binomial(link = "logit"))

#look at model summary 
summary(thick_glm)

#obtain estimated means (probabilities of score=1 by lobe)
thick_emm <- emmeans(thick_glm, ~ lobe, type = "response")

#see the pairwise comparisons
pairs(thick_emm)
```

Significant findings:
*RUL - RML: OR = 10.4 => RUL's odds of disease is 10.4 times that of RML
*RUL - RLL: OR = 2.48 => RUL's odds of disease is 2.48 times that of RLL
*RUL - LUS: OR = 4.07 => RUL's odds of disease is 4.07 times that of LUS
*RUL - LLS: OR = 10.4 => RUL's odds of disease is 10.4 times that of LLS
*RUL - LLL: OR = 3.68 => RUL's odds of disease is 3.68 times that of LLL

## ICC Calculations

### With Latent Variance

```{r}
#define function
ICC_f <- function(T1, T2, VarT1, VarT2, CovT1T2, c=(pi^2)/3){
  denom <- (T1+T2+c)
  estimate <- T1/(denom)
  partial1 <- (T2+c)/(denom^2)
  partial2 <- (-T1)/(denom^2)
  
  var_est <- (partial1^2 * VarT1) + (partial2^2 * VarT2) + (2*partial1*partial2*CovT1T2)
  se_est <- sqrt(var_est)
  
  ci_lower <- estimate - (1.96*se_est)
  ci_upper <- estimate + (1.96*se_est)
  
  #print estimates
  print(estimate)
  print(ci_lower)
  print(ci_upper)
}

print('Tree-In-Bud')
ICC_f(T1=3.3175, T2=0.1353, VarT1=0.2166, VarT2=0.01077, CovT1T2=0.001336)

print('')
print('GGO')
ICC_f(T1=5.0430, T2=0.5583, VarT1=1.3298, VarT2=0.09725, CovT1T2=0.02422)

print('')
print('Consolidation')
ICC_f(T1=2.2156, T2=0.3958, VarT1=0.1566, VarT2=0.03091, CovT1T2=-0.00315)

print('')
print('Bronchiectasis')
ICC_f(T1=4.1952, T2=0.9182, VarT1=0.3929, VarT2=0.04153, CovT1T2=-0.00076)
```

### Without Latent Variance

```{r}
#define function
ICC_f <- function(T1, T2, VarT1, VarT2, CovT1T2){
  denom <- (T1+T2)
  estimate <- T1/(denom)
  partial1 <- (T2)/(denom^2)
  partial2 <- (-T1)/(denom^2)
  
  var_est <- (partial1^2 * VarT1) + (partial2^2 * VarT2) + (2*partial1*partial2*CovT1T2)
  se_est <- sqrt(var_est)
  
  ci_lower <- estimate - (1.96*se_est)
  ci_upper <- estimate + (1.96*se_est)
  
  #print estimates
  print(estimate)
  print(ci_lower)
  print(ci_upper)
}

print('Tree-In-Bud')
ICC_f(T1=3.3175, T2=0.1353, VarT1=0.2166, VarT2=0.01077, CovT1T2=0.001336)

print('')
print('GGO')
ICC_f(T1=5.0430, T2=0.5583, VarT1=1.3298, VarT2=0.09725, CovT1T2=0.02422)

print('')
print('Consolidation')
ICC_f(T1=2.2156, T2=0.3958, VarT1=0.1566, VarT2=0.03091, CovT1T2=-0.00315)

print('')
print('Bronchiectasis')
ICC_f(T1=4.1952, T2=0.9182, VarT1=0.3929, VarT2=0.04153, CovT1T2=-0.00076)
```